Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-03-15 introduction/Subsections/Introduction.tex
Updates from Overleaf

Music is an essential part of human culture, history, and society. It has existed in our lives since ancient times and has evolved, reflecting changes in our cultures, traditions, and beliefs. Despite its importance, the study of music has always been a challenging and complex task due to its subjective nature and the vast array of musical styles, genres, and cultures that exist worldwide.

Humanity has been studying music to understand its underlying structure for millennia. From the ancient Greeks, who believed that music had a mathematical and cosmic harmony, to the early works of theorists such as Rameau, Fux, and Schenker, and more recent approaches such as generative theory and corpus studies, music has been a subject of fascination and inquiry. This rich music theory and analysis history provides a valuable foundation for guiding and evaluating MIR algorithms and models.

@@

Let's pick a core MSA task as a sample to display such complexity: Music Structure Analysis (MSA) is an interdisciplinary field that aims to understand the structure of music. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges such as boundary placement ambiguity and similarity quantification. The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity.

The checkerboard kernel technique is a simple and effective method for Music Structure Analysis (MSA) based on the homogeneity principle. The kernel, with a checkerboard-like structure, is convolved over the main diagonal of a Self-Similarity Matrix (SSM) such as $S_{ij}$ is the similarity between time frames $i$ and $j$, $\vec{x}_i$ is the vector representation of time frame $i$, and $\left| \vec{x}_i \right|$ is the Euclidean norm of vector $\vec{x}_i$. The dot product of two vectors $\vec{x}_i$ and $\vec{x}_j$ is denoted by $\vec{x}_i \cdot \vec{x}_j$.

\input{introduction/subsections/equations/SSM equation}

This yields a novelty curve that highlights sudden changes in the selected musical features (e.g. chroma, MFCCs, beat-synchronous features, onset and offset detection, and harmonic and percussive separation) from which to extract the segment boundaries.

\input{introduction/subsections/equations/Novelty equation}

Mathematical operations are used on predefined principles such as homogeneity and repetition to analyze a musical structure. It relies on analyzing features extracted from a musical signal and identifying peaks in a novelty curve to extract segment boundaries. In contrast, human aural skills rely on the perception and interpretation of music by human listeners based on cultural and academical background and can capture subtle nuances of musical structure, but those still are subjective and can vary across listeners.

Combining technical tasks, aural skills, and music perception is challenging due to the complexity and subjectivity of musical perception, discrepancies between the features extracted by DSP techniques and the features perceived by humans, and the variability of musical signals and individual differences in musical perception. DSP techniques may not always reflect the way humans perceive music, and may not capture the subtle variations in timing or harmonic relationships that are critical to musical perception. Variability in musical signals due to instrumentation, genre, and performance style further complicates the development of DSP techniques that can be generalized to different musical contexts.

>>>>

Music is an essential part of human culture, history, and society. It has existed since ancient times and evolved, reflecting changes in our cultures, traditions, and beliefs. Despite its importance, the study of music has always been a challenging and complex task due to its subjective nature and the vast array of musical styles, genres, and cultures that exist worldwide.

Humanity has been studying music to understand its underlying structure for millennia. From the ancient Greeks, who believed that music had a mathematical and cosmic harmony, to the early works of theorists such as Rameau, Fux, and Schenker, and more recent approaches such as generative theory and corpus studies, music has been a subject of fascination and inquiry. This rich music theory and analysis history provide a valuable foundation for guiding and evaluating MIR algorithms and models.

@@

### 2023-03-15 introduction/Subsections/Motivation.tex
Updates from Overleaf

As a music lover, professional musician, and connoisseur, I am interested in gaining a deeper understanding of the fundamental building of musical composition, the various techniques used to create musical works, and capturing underlying relationships among them. I am deeply interested in learning how to retrieve embedded information within music. The study of music structure analysis provides a means to systematically analyze and understand musical pieces. It has traditionally been approached through music theory, formal musical notation, musical analysis techniques, and machine learning, which I plan to blend.

With the advent of Artificial Intelligence (AI) and Machine Learning (ML), Music Information Retrieval (MIR) has expanded to encompass musically motivated neural network architectures. These networks are designed to incorporate musical concepts and perceptual approaches and use this knowledge to analyze and understand the structure of musical pieces in a way that goes beyond traditional music analysis techniques.

As a professional musician and music transcriber, I am interested in understanding how Deep Neural Networks (DNNs) filter musical features. In other words, I am eager to learn how DNNs listen to music.

By studying music structure analysis using musically motivated neural networks, we will gain an in-depth understanding of the latest advancements in the field and how these techniques are being used to revolutionize our understanding of music. Additionally, we will be able to contribute to developing new and innovative methods for musical analysis.

>>>>

We argue that MIR research needs to incorporate a more balanced approach that considers the interdisciplinary nature of music and the importance of other domains beyond DSP.

Let's pick a core MSA task as a sample to display such complexity: Music Structure Analysis (MSA) is an interdisciplinary field that aims to understand the structure of music. However, due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification. The main principles of MSA were initially defined as homogeneity, novelty, and repetition, with the addition of regularity.

The checkerboard kernel technique is a simple and effective method for Music Structure Analysis (MSA) based on the homogeneity principle. The kernel, with a checkerboard-like structure, is convolved over the main diagonal of a Self-Similarity Matrix (SSM) such as $S_{ij}$ is the similarity between time frames $i$ and $j$, $\vec{x}_i$ is the vector representation of time frame $i$, and $\left| \vec{x}_i \right|$ is the Euclidean norm of vector $\vec{x}_i$. The dot product of two vectors $\vec{x}_i$ and $\vec{x}_j$ is denoted by $\vec{x}_i \cdot \vec{x}_j$.

\input{introduction/subsections/equations/SSM equation}

This yields a novelty curve that highlights sudden changes in the selected musical features (some examples would be chroma, MFCCs, beat-synchronous features, onset and offset detection, and harmonic and percussive separation) from which to extract the segment boundaries.

\input{introduction/subsections/equations/Novelty equation}

Mathematical operations are used on predefined principles such as homogeneity and repetition to analyze a musical structure. It analyzes features extracted from a musical signal and identifies peaks in a novelty curve to extract segment boundaries. In contrast, human aural skills rely on the perception and interpretation of music by human listeners based on cultural and academic background. They can capture subtle nuances of musical structure, but those still are subjective and can vary across listeners.

Combining technical tasks, aural skills, and music perception is challenging due to the complexity and subjectivity of musical perception, discrepancies between the features extracted by DSP techniques and the features perceived by humans, and the variability of musical signals and individual differences in musical perception. DSP techniques may not always reflect how humans perceive music and may not capture the subtle variations in timing or harmonic relationships critical to musical perception. Variability in musical signals due to instrumentation, genre, and performance style further complicates the development of DSP techniques that can be generalized to different musical contexts.

As a music lover, professional musician, and connoisseur, I am interested in gaining a deeper understanding of the fundamental building of musical composition, the various techniques used to create musical works, and capturing underlying relationships among them using DN.  I am eager to learn how DNNs listen to music. I am deeply interested in learning how to retrieve embedded information within music.

### 2023-03-15 introduction/Subsections/Objectives.tex
Updates from Overleaf

>>>>

With the advent of Machine Learning (ML), MIR has expanded to encompass musically motivated neural network architectures. These networks are designed to incorporate musical concepts and perceptual approaches and use this knowledge to analyze and understand the structure of musical pieces in a way that goes beyond traditional music analysis techniques.

By studying music structure analysis using musically motivated neural networks, we will gain an in-depth understanding of the latest advancements in the field and how these techniques are being used to revolutionize our understanding of music. Additionally, we will be able to contribute to developing new and innovative methods for musical analysis.

### 2023-03-15 introduction/subsections/underlying language vs waveform.tex
Updates from Overleaf

The underlying composition represents the musical and lyrical elements of the song in a fixed and notated form. It includes the melody, rhythm, and harmony of the song as well as the accompanying lyrics. The underlying composition is the intellectual property of the songwriters, composers, musicians, or publishers, who have exclusive rights to reproduce, distribute, and perform the work.

The master recording, on the other hand, is the tangible recording of the sound of the song. It captures the performance, production, and engineering aspects of the work. The master recording is the intellectual property of the performers, producers, or sound engineers, who have exclusive rights to reproduce, distribute, and prepare derivative works of the sound recording.

In essence, the underlying composition is a fixed and notated representation of the song's musical and lyrical elements, while the master recording is a tangible recording of the performance, production, and engineering aspects of the song. The underlying composition is owned by the songwriters, composers, musicians, or publishers, while the master recording is owned by the performers, producers, or sound engineers.

>>>>

The underlying composition represents the musical and lyrical elements of the song in a fixed, heuristic, and possibly notated form. It includes the melody, rhythm, harmony of the song, and accompanying lyrics, if any. In the lawsuit domain, the underlying composition is the intellectual property of the songwriters, composers, musicians, or publishers, who have exclusive rights to reproduce, distribute, and perform the work.

The master recording, on the other hand, is the tangible recording of the song's sound. It captures the performance, production, and engineering aspects of the work. The master recording is the intellectual property of the performers, producers, or sound engineers, who have exclusive rights to reproduce, distribute, and prepare derivative works of the sound recording.

The underlying composition is a fixed and notated representation of the song's musical and lyrical elements. At the same time, the master recording is a tangible recording of the song's performance, production, and engineering aspects. The underlying composition is owned by the songwriters, composers, musicians, or publishers, while the master recording is owned by the performers, producers, or sound engineers.

### 2023-03-15 methods/subsections/test.tex
Updates from Overleaf

%Single
\input{Figures/Figures CNN/Single Neuron}
%NN
\input{Figures/Figures CNN/Network_graph}

%Backpropagation
\input{Figures/Figures CNN/Multi-layer_NN}

%CNN
A high-level illustration of a simple convolutional neural network indicating convolutional layers, pooling layers, and fully-connected layers without details (number of channels or neurons per layer or the input image size) can be seen in XXX.
\input{Figures/Figures CNN/CNN_diagram}

%Single CNN
\input{Figures/Figures CNN/Single_CNN}

%Single pooling layer
\input{Figures/Figures CNN/Single_Pooling_Layer}

%Backpropagation
\input{Figures/Figures CNN/Back_propagation}

>>>>

%Single
\input{figures/Figures CNN/Single Neuron}
%NN
\input{figures/Figures CNN/Network_graph}

%Backpropagation
\input{figures/Figures CNN/Multi-layer_NN}

%CNN
A high-level illustration of a simple convolutional neural network indicating convolutional layers, pooling layers, and fully-connected layers without details on the number of channels or neurons per layer or the input data size can be seen in XXX.
\input{figures/Figures CNN/CNN_diagram}

%Single CNN
\input{figures/Figures CNN/Single_CNN}

%Single pooling layer
\input{figures/Figures CNN/Single_Pooling_Layer}

%Backpropagation
\input{figures/Figures CNN/Back_propagation}
